---
title: "final_project_AA"
author: "lucy"
date: "December 5, 2018"
output:
  html_document: default
  pdf_document: default
---

Load packages
```{r}
### Load helper packages ###
loadlibs = function(libs) {
  for(lib in libs) {
    class(lib)
    if(!do.call(require,as.list(lib))) {install.packages(lib)}
    do.call(require,as.list(lib))
  }
}
libs = c("tidyr","magrittr","purrr","dplyr","stringr","readr","data.table", "lubridate")
loadlibs(libs)
```

Load data 2007-2008,2005-2006 from RNHANES
```{r}
library(RNHANES)
#Df_total <- nhanes_data_files() # gives you the list of files.

#demographic, gender, age, race, total number of people in the family and household
demo07 <- nhanes_load_data("DEMO_E", year = "2007-2008")%>%data.frame()%>%as_tibble()%>%select(SEQN, RIAGENDR, DMDHRAGE,RIDRETH1, DMDFMSIZ,DMDHHSIZ)
demo06 <- nhanes_load_data("DEMO_D", year = "2005-2006")%>%data.frame()%>%as_tibble()%>%select(SEQN, RIAGENDR, DMDHRAGE,RIDRETH1, DMDFMSIZ,DMDHHSIZ)

#Ophthalmology, Y(retinopothy), retinal microaneurysms, retinal hard exudate
reti07 = nhanes_load_data("OPXRET_E", year = "2007-2008")%>%data.frame()%>%as_tibble()%>%select(SEQN,OPDDARMA,OPDSARMA,OPDUHEM,OPDUHE,OPDUSE)
reti06 = nhanes_load_data("OPXRET_D", year = "2005-2006")%>%data.frame()%>%as_tibble()%>%select(SEQN,OPDDARMA,OPDSARMA,OPDUHEM,OPDUHE,OPDUSE)


#vision,had surgery for cataract before, visual acuity of left eyes and right eyes, wearing glasses or not
vision_07 <- nhanes_load_data("VIX_E", year = "2007-2008")%>%data.frame()%>%as_tibble%>%select(SEQN,VIQ200,VIDLOVA, VIDROVA, VIQ150)
vision_06 <- nhanes_load_data("VIX_D", year = "2005-2006")%>%data.frame()%>%as_tibble%>%select(SEQN,VIQ200,VIDLOVA, VIDROVA,VIQ150)


#bmx, BOdy mass index, weight
bmx_07 <- nhanes_load_data("BMX_E", year = "2007-2008")%>%data.frame()%>%as_tibble%>%select(SEQN,BMXBMI, BMXWT)
bmx_06 <- nhanes_load_data("BMX_D", year = "2005-2006")%>%data.frame()%>%as_tibble%>%select(SEQN,BMXBMI, BMXWT)

#glu, fasting glucose
glu_07 <- nhanes_load_data("GLU_E", year = "2007-2008")%>%data.frame()%>%as_tibble%>%select(SEQN,LBXGLU)
glu_06 <- nhanes_load_data("GLU_D", year = "2005-2006")%>%data.frame()%>%as_tibble%>%select(SEQN,LBXGLU)


```

Join data, Split data into train and test
```{r}

#Join the data in 2007
patient07 = reti07%>%left_join(demo07,by = "SEQN")%>%left_join(vision_07,by = "SEQN")%>%left_join(bmx_07,by = "SEQN")%>%left_join(glu_07,by = "SEQN")

#create Y column
patient07=patient07%>%mutate(RETI = ifelse(OPDDARMA==1|OPDSARMA==1,1,0))%>%select(-OPDDARMA,-OPDSARMA,-SEQN)

#join data in 2006
patient06 = reti06%>%left_join(demo06,by = "SEQN")%>%left_join(vision_06,by = "SEQN")%>%left_join(bmx_06,by = "SEQN")%>%left_join(glu_06,by = "SEQN")

patient06=patient06%>%mutate(RETI = ifelse(OPDDARMA==1|OPDSARMA==1,1,0))%>%select(-OPDDARMA,-OPDSARMA,-SEQN)

#combine all the data
patient = patient07%>%rbind(patient06)

patient=patient%>%filter(!is.na(RETI))%>%filter(DMDHRAGE>=40)


#randominze the data and split 
set.seed(504)
patient=patient[sample(1:nrow(patient)),]

summary(patient)

#70% as train, 30% as test
p_train = patient[1:as.integer(nrow(patient)*0.7),]
p_test = patient[-(1:as.integer(nrow(patient)*0.7)),]


```


Impute training data and include all imputed data as training data
Impute testing data and include all imputed data as testing data

```{r}
#imputation
#I use mutiple imputation to impute the missing data. Because in this case, the missing data is MCAR so that can use MICE to impute.
#install.packages("mice")
library(mice) 

ptrain = mice(p_train %>% 
              select(-RETI) %>% 
              mutate_if(is.character, as.factor),m=10,maxit = 5) 

# Get all imputed data set
alldata = function(x,y){
  imputed_p = NULL
  for (i in (1:5)){
    imp = complete(x,i) %>% as_tibble()
    names(imp) = lapply(names(imp), paste0, "_imputed")
    # Rename those columns in case you want to column append
    miss = apply(y,FUN = (function(x) any(is.na(x))),MARGIN = 2)
    miss = miss[-which(names(miss)=="RETI")]
    imp = imp[,paste0(names(miss %>% t() %>% data.frame()),"_imputed")]
    imputed = data.frame(RETI = y$RETI) %>% bind_cols(imp) %>% as_tibble()
    imputed_p = imputed_p%>%rbind(imputed)
  }
  imputed_p
    
}

imputed_ptrain = alldata(ptrain,p_train)


#get test imputed data
ptest = mice(p_test %>% 
              select(-RETI) %>% 
              mutate_if(is.character, as.factor),m=10,maxit = 5) 
imputed_ptest = alldata(ptest,p_test)

```

Without ophthalmology variables:  this analysis is intended for the situation where you know very little about the person (e.g. demographics only), and you want to know what effect measuring fasting blood glucose will have on risk of retinopathy.
Preparation: Data that has glucose or not
```{r}
p_train_noo_noglu = imputed_ptrain%>%select(-c(OPDUHEM_imputed,OPDUHE_imputed,OPDUSE_imputed,LBXGLU_imputed))
p_train_noo_glu = imputed_ptrain%>%select(-c(OPDUHEM_imputed,OPDUHE_imputed,OPDUSE_imputed))


p_test_noo_noglu = imputed_ptest%>%select(-c(OPDUHEM_imputed,OPDUHE_imputed,OPDUSE_imputed,LBXGLU_imputed))
p_test_noo_glu = imputed_ptest%>%select(-c(OPDUHEM_imputed,OPDUHE_imputed,OPDUSE_imputed))

```

trian the model
```{r}
#logistic regression
ptrain_lr_noo_noglu = glm(RETI == 1 ~ ., data =p_train_noo_noglu ,  family = binomial("logit"))

ptrain_lr_noo_glu = glm(RETI == 1 ~ ., data =p_train_noo_glu ,  family = binomial("logit"))

summary(ptrain_lr_noo_noglu)
summary(ptrain_lr_noo_glu)

```

Test model and See the comparation results of -ophthalmology,(+-glucose)
```{r}
library(pROC)

#par(mar=c(5,4,5,3), xpd=TRUE)
#lr roc
result_lr_noo_noglu = p_test_noo_noglu%>%select(RETI)%>%bind_cols(pred = predict(ptrain_lr_noo_noglu, p_test_noo_noglu, type  = "response"))%>%na.omit()
result_lr_noo_glu = p_test_noo_glu%>%select(RETI)%>%bind_cols(pred = predict(ptrain_lr_noo_glu,p_test_noo_glu, type  = "response"))%>%na.omit()


lrroc_noo_noglu <- roc(result_lr_noo_noglu$RETI,result_lr_noo_noglu$pred)
plot(lrroc_noo_noglu, print.auc=TRUE,col = "red",print.auc.x = 0.2,print.auc.y = 0.5)

par(new=TRUE)

lrroc_noo_glu <- roc(result_lr_noo_glu$RETI,result_lr_noo_glu$pred)
plot(lrroc_noo_glu, print.auc=TRUE,col = "green",print.auc.x = 0.4,print.auc.y = 0.6)


#imputed_ptest$RETI = as.factor(imputed_ptest$RETI)
#rf roc
#result_rf = imputed_ptest%>%select(RETI)%>%bind_cols(pred = predict(pforest, imputed_ptest, type = "prob"))%>%na.omit()

#rfroc <- roc(result_rf$RETI,as.numeric(result_rf$pred))
#plot(rfroc, print.auc=TRUE,col = "green",add=TRUE,print.auc.x = 0.4,print.auc.y = 0.4)


legend("topright", c("Without ophthalmology Without Glucose", "Without ophthalmology With Glucose"), lty=2, 
    col = c("red", "green"), bty="n", inset=c(0,0.5))

#without glucose
glm.pred1 = rep("0",7730)
glm.pred1[result_lr_noo_noglu$pred>.5] = "1"
table(glm.pred1,p_test_noo_noglu$RETI)

#with glucose
glm.pred2 = rep("0",7730)
glm.pred2[result_lr_noo_glu$pred>.5] = "1"
table(glm.pred2,p_test_noo_glu$RETI)

```


With ophthalmology variables:  this analysis is intended for the situation where you know very little about the person (e.g. demographics only), and you want to know what effect measuring fasting blood glucose will have on risk of retinopathy.
Preparation: Data that has glucose or not
```{r}
p_train_o_noglu = imputed_ptrain%>%select(-LBXGLU_imputed)
p_train_o_glu = imputed_ptrain


p_test_o_noglu = imputed_ptest%>%select(-LBXGLU_imputed)
p_test_o_glu = imputed_ptest

```

trian the model
```{r}
#logistic regression
ptrain_lr_o_noglu = glm(RETI == 1 ~ ., data =p_train_o_noglu ,  family = binomial("logit"))

ptrain_lr_o_glu = glm(RETI == 1 ~ ., data =p_train_o_glu ,  family = binomial("logit"))

summary(ptrain_lr_o_noglu)
summary(ptrain_lr_o_glu)

```

Test model and See the comparation results of +ophthalmology,(+-glucose)
```{r}
library(pROC)

par(mar=c(5,4,5,3), xpd=TRUE)
#lr roc
result_lr_o_noglu = p_test_o_noglu%>%select(RETI)%>%bind_cols(pred = predict(ptrain_lr_o_noglu, p_test_o_noglu, type  = "response"))%>%na.omit()
result_lr_o_glu = p_test_o_glu%>%select(RETI)%>%bind_cols(pred = predict(ptrain_lr_o_glu,p_test_o_glu, type  = "response"))%>%na.omit()


lrroc_o_noglu <- roc(result_lr_o_noglu$RETI,result_lr_o_noglu$pred)
plot(lrroc_o_noglu, print.auc=TRUE,col = "red",print.auc.x = 0.2,print.auc.y = 0.5)

par(new=TRUE)

lrroc_o_glu <- roc(result_lr_o_glu$RETI,result_lr_o_glu$pred)
plot(lrroc_o_glu, print.auc=TRUE,col = "green",print.auc.x = 0.4,print.auc.y = 0.6)


legend("topright", c("With ophthalmology Without Glucose", "With ophthalmology With Glucose"), lty=2, 
    col = c("red", "green"), bty="n", inset=c(0,0.5))

#without glucose
glm.pred3 = rep("0",7730)
glm.pred3[result_lr_o_noglu$pred>.5] = "1"
table(glm.pred3,p_test_o_noglu$RETI)

#with glucose
glm.pred4 = rep("0",7730)
glm.pred4[result_lr_o_glu$pred>.5] = "1"
table(glm.pred4,p_test_o_glu$RETI)

```


